Detection and partial characterisation of stellar multiplicity with Gaia

### 2018-07-02 article/ms.tex
stuff

\begin{figure*}
	\includegraphics[width=1.0\textwidth]{../figures/todo.png}
    \caption{Fraction of \Gaia\ sources with reported radial velocities
		     as a function of source properties. Within the adopted source
		     parameter ranges (gray; see Section \ref{sec:sb2_methods})
		     the radial velocity completeness is approximately flat.
		     We flag sources within this range that do not have reported 
		     radial velocities to be candidate SB2 systems.}
    \label{fig:sb2_rvs_completeness}
\end{figure*}

\begin{figure*}
	\includegraphics[width=1.0\textwidth,height=3.0in]{../figures/todo.png}
    \caption{Distribution of 
    			(a) photometric \bp\ - \rp\ excess factor, 
				(b) astrometric unit weight error, 
					and 
				(c) photometric variability
			of candidate SB2 type systems relative to a control sample.}
    \label{fig:sb2_histograms}
\end{figure*}

\begin{figure*}
	\includegraphics[width=1.0\textwidth,height=3.0in]{../figures/todo.png}
    \caption{Hertzsprung-Russell diagram of X \Gaia\ sources that are in the
    		 ranges specified in Equation \ref{eq:rvs_parameter_ranges} (gray).
		 	 Overlaid in colour we show the SB2 candidates we deduce by the
			 absence of reported radial velocity.}
    \label{fig:sb2_hrd}
\end{figure*}

@@

Many higher-order star systems are unresolved point sources in \Gaia, but the
stellar companion(s) may have sufficient mass to measurably affect the tangential 
velocity of the primary and produce a detectable jitter in the astrometric 
positions. There are many other sources of astrometric jitter, but here we show 
that the principle effect of astrometric excess noise is due to stellar 
companions. Note that due to the astrometry `degree of freedom bug' in the 
processing of the second \Gaia\ data release \citep[e.g., see Appendix A1 of ][]{Lindegren:2018}, 
that resulted in 80\% of sources having zero astrometric excess noise,
we instead make use of the astrometric unit weight error $u$
\begin{eqnarray}
	u = \sqrt{\frac{\chi^2_{al}}{N_{obs,al} - 5}}
\end{eqnarray}
\noindent{}where $\chi^2_{al}$ is the astrometric $\chi^2$ value
in the along-scan direction (column \texttt{astrometric\_chi2\_al}) and 
$N_{obs,al}$ is the number of astrometric observations in the along-scan
direction (column \texttt{astrometric\_n\_obs\_al}). The values of this
quantity can be interpreted similarly to how the astrometric excess noise
would be interpreted (e.g., large values are less consistent with a
single star model), although like the radial velocity and astrometric
jitter, it must be considered in context with stars of similar colour,
apparent magnitude (and preferably, absolute magnitude).

@@

>>>>

\begin{figure*}
	\includegraphics[width=1.0\textwidth]{../figures/sb2_rvs_completeness.pdf}
    \caption{Fraction of \Gaia\ sources with reported radial velocities
		     as a function of source properties. Within the adopted source
		     parameter ranges (gray; see Section \ref{sec:sb2_methods})
		     the radial velocity completeness is approximately flat.
		     We flag sources within this range that do not have reported 
		     radial velocities to be candidate SB2 systems.}
    \label{fig:sb2_rvs_completeness}
\end{figure*}

\begin{figure}
	\includegraphics[width=0.4\textwidth]{../figures/sb2_histograms.pdf}
    \caption{Distribution of 
    			(a) astrometric unit weight error (see Eq. \ref{eq:astrometric_unit_weight_error}),
			%$u = \left(\chi_{al}^2/(N_{al} - 5)\right)^{1/2}$
			%		where $\chi_{al}^2$ and $N_{al}$ is the astrometric goodness-of-fit and number 
			%		of observations in the along-scan direction, respectively (columns 
			%			\texttt{astrometric\_chi2\_al} and \texttt{astrometric\_n\_obs\_al}), 
				(b) photometric \bp\ - \rp\ excess factor, 				
	 				and 
				(c) photometric \rp\ variability (see Eq. \ref{eq:photometric_variability})
			of candidate SB2 type systems relative to a control sample of the same size.}
    \label{fig:sb2_histograms}
\end{figure}

\begin{figure*}
	\includegraphics[width=1.0\textwidth]{../figures/sb2_sky_structure.pdf}
    \caption{Fraction of sources (per sky bin) without reported radial velocities
    		 for all sources with $G \lesssim 13$ (top) and sources in the ranges
		 	 specified by Eqs \ref{eq:sb2_ranges} (bottom), where we assert that a missing
			 radial velocity measurement signifies a likely SB2 candidate . The color scale is 
			 arbitrarily set to highlight structure, where black indicates a higher
			 fraction of sources do not have reported radial velocities. Crowding
			 in the galactic plane likely results in some sources not having radial
			 velocities reported. The large scale structure visible in both axes is
			 a combined effect of the initial \Gaia\ source list, the scanning law,
			 and star forming regions (i.e., where emission in the \ion{Ca}{2} 
			 triplet likely causes issues for radial velocity determination).}
    \label{fig:sb2_sky_structure}
\end{figure*}

@@

Many higher-order star systems are unresolved point sources in \Gaia, but the
stellar companion(s) may have sufficient mass to measurably affect the tangential 
velocity of the primary and produce a detectable jitter in the astrometric 
positions. There are many other sources of astrometric jitter, but here we show 
that the principle effect of astrometric excess noise is due to stellar 
companions. Note that due to the astrometry `degree of freedom bug' in the 
processing of the second \Gaia\ data release \citep[e.g., see Appendix A1 of ][]{Lindegren:2018}, 
that resulted in 80\% of sources having zero astrometric excess noise,
we instead make use of the astrometric unit weight error $u$
\begin{eqnarray}
	u = \sqrt{\frac{\chi^2_{al}}{N_{obs,al} - 5}} \label{eq:astrometric_unit_weight_error}
\end{eqnarray}
\noindent{}where $\chi^2_{al}$ is the astrometric $\chi^2$ value
in the along-scan direction (column \texttt{astrometric\_chi2\_al}) and 
$N_{obs,al}$ is the number of astrometric observations in the along-scan
direction (column \texttt{astrometric\_n\_obs\_al}). The values of this
quantity can be interpreted similarly to how the astrometric excess noise
would be interpreted (e.g., large values are less consistent with a
single star model), although like the radial velocity and astrometric
jitter, it must be considered in context with stars of similar colour,
apparent magnitude (and preferably, absolute magnitude).

@@

\todo{bp rp excess factor}

\begin{eqnarray}
	\textrm{Amp} = \log_{10}\left(\sqrt{N_{obs,al}}\frac{\sigma_{\langle{}f_G\rangle}}{\langle{}f_G\rangle}\right) \label{eq:photometric_variability}
\end{eqnarray}
